---
title: "Data Exploration"
format: html
editor: visual
---

## Data Exploration

Load the packages. If not installed please install first.

Overview of data

```{r}
library(foreign)
library(dplyr)


W01 <- read.spss("../00.Data/01. covid-19_wave1_survey_cls.sav", to.data.frame = TRUE)
summary(W01)
head(W01)

W02 <- read.spss("../00.Data/02. covid-19_wave2_survey_cls.sav", to.data.frame = TRUE)
summary(W02)
head(W02)

W03 <- read.spss("../00.Data/03. covid-19_wave3_survey_cls.sav", to.data.frame = TRUE)
summary(W03)
head(W03)

```

### Select Relevant Rows

The study looks into the participants with long-standing illness. Hence the following variable will be selected.

\*Demo = Demographic information Acronym

\*Mod\_ = Moderator

## Wave 1

**Wave 1** Variable Table

| Type | Variable Acronym | Explanation |
|---------------|---------------|------------------------------------------|
| Demo | CW1_PSEX | Sex of Respondent |
| Mod_Cov | CW1_COVIDSYMPT_1 | Past 2 weeks COVID19 symptoms: Fever |
| Mod_Cov | CW1_COVIDSYMPT_2 | Past 2 weeks COVID19 symptoms: Cough - dry |
| Mod_Cov | CW1_COVIDSYMPT_3 | Past 2 weeks COVID19 symptoms: Cough - mucus or phlegm |
| Mod_Cov | CW1_COVIDSYMPT_4 | Past 2 weeks COVID19 symptoms: Sore throat |
| Mod_Cov | CW1_COVIDSYMPT_5 | Past 2 weeks COVID19 symptoms: Chest tightness |
| Mod_Cov | CW1_COVIDSYMPT_6 | Past 2 weeks COVID19 symptoms: Shortness of breath |
| Mod_Cov | CW1_COVIDSYMPT_7 | Past 2 weeks COVID19 symptoms: Runny nose |
| Mod_Cov | CW1_COVIDSYMPT_8 | Past 2 weeks COVID19 symptoms: Nasal congestion |
| Mod_Cov | CW1_COVIDSYMPT_18 | Past 2 weeks COVID19 symptoms: Sneezing |
| Mod_Cov | CW1_COVIDSYMPT_10 | Past 2 weeks COVID19 symptoms: Muscle or body aches |
| Mod_Cov | CW1_COVIDSYMPT_11 | Past 2 weeks COVID19 symptoms: Fatigue |
| Mod_Cov | CW1_COVIDSYMPT_12 | Past 2 weeks COVID19 symptoms: Unusual loose motions or diarrhoea |
| Mod_Cov | CW1_COVIDSYMPT_16 | Past 2 weeks COVID19 symptoms: Vomiting |
| Mod_Cov | CW1_COVIDSYMPT_13 | Past 2 weeks COVID19 symptoms: Loss of smell |
| Mod_Cov | CW1_COVIDSYMPT_14 | Past 2 weeks COVID19 symptoms: Loss of taste |
| Mod_Cov | CW1_COVIDSYMPT_17 | Past 2 weeks COVID19 symptoms: Skin rash |
| Mod_Cov | CW1_COVIDSYMPT_19 | Past 2 weeks COVID19 symptoms: Headaches |
| Mod_Cov | CW1_COVIDSYMPT_20 | Past 2 weeks COVID19 symptoms: Other |
| Mod_Cov | CW1_COVIDSYMPT_23 | Past 2 weeks no COVID19 symptoms experienced |
| Mod_Health | CW1_GHQ | Post-C19: Respondent self-assessment of general health |
| Mod_Health | CW1_GHQPRECOVID | Pre-C19: Respondent self-assessment of general health in 3 months preceding outbreak |
| IV | CW1_LLI_1 | long-standing illness 1: Cancer |
|  | CW1_LLI_2 | long-standing illness: Cystic fibrosis |
|  | CW1_LLI_3 | long-standing illness: Asthma |
|  | CW1_LLI_4 | long-standing illness: Chronic Obstructive Pulmonary Disease |
|  | CW1_LLI_5 | long-standing illness: Wheezy bronchitis |
|  | CW1_LLI_6 | long-standing illness: Diabetes |
|  | CW1_LLI_7 | long-standing illness: Recurrent backache/prolapsed disc/sciatica/other back probm |
|  | CW1_LLI_8 | long-standing illness: Problems with hearing |
|  | CW1_LLI_9 | long-standing illness: High blood pressure |
|  | CW1_LLI_10 | long-standing illness: Heart disease, congenital or acquired |
|  | CW1_LLI_11 | long-standing illness: Depression or other emotional/nervous/psychiatric probs |
|  | CW1_LLI_12 | long-standing illness: Obesity |
|  | CW1_LLI_13 | long-standing illness: Chronic Obstructive Airways Disease |
|  | CW1_LLI_14 | long-standing illness: Infection |
|  | CW1_LLI_15 | long-standing illness: HIV/Immunodeficiency |
|  | CW1_LLI_16 | long-standing illness: Conditions affecting the brain and nerves |
|  | CW1_LLI_17 | Respondent suffering from a long-standing illness: |
| Mod_Lonely | CW1_LONELY_1 | How often - lack companionship |
|  | CW1_LONELY_2 | How often - left out |
|  | CW1_LONELY_3 | How often - isolated from others |
|  | CW1_LONELY_4 | How often - lonely |
| DV_Kessler6 | CW1_PHDE | Felt depressed |
|  | CW1_PHHO | felt hopeless |
|  | CW1_PHRF | felt restless |
|  | CW1_PHEE | felt everything was an effort |
|  | CW1_PHWO | felt worthless |
|  | CW1_PHNE | felt nervous |
|  | CW1_HARM | has harmed self on purpose |
| DV_WEMWBS | CW1_WEMWBS_1 | feeling optimistic about the future |
|  | CW1_WEMWBS_2 | feeling useful |
|  | CW1_WEMWBS_3 | feeling relaxed |
|  | CW1_WEMWBS_4 | dealing with problems well |
|  | CW1_WEMWBS_5 | thinking clearly |
|  | CW1_WEMWBS_6 | thinking close to other people |
|  | CW1_WEMWBS_7 | able to make up my own mind about things |
| DV_MALAISE | CW3_MALAISE_1 | feel tired most of the time |
|  | CW3_MALAISE_2 | feel miserable or depressed |
|  | CW3_MALAISE_3 | get worried about things |
|  | CW3_MALAISE_4 | get in a violent rage |
|  | CW3_MALAISE_5 | suddenly become scared for no good reason |
|  | CW3_MALAISE_6 | easily upset or irritated |
|  | CW3_MALAISE_7 | constantly keyed up and jittery |
|  | CW3_MALAISE_8 | every little thing get on your nerves and wear you out |
|  | CW3_MALAISE_9 | heart often race like mad |
| DV_GHQ | CW1_GHQ121 | able to concentrate |
|  | CW1_GHQ122 | lost much sleep over worry |
|  | CW1_GHQ123 | felt that playing a useful part in things |
|  | CW1_GHQ124 | felt capable of making decisions about things |
|  | CW1_GHQ125 | felt constantly under strain |
|  | CW1_GHQ126 | couldn’t overcome difficulties |
|  | CW1_GHQ127 | able to enjoy normal day to day activities |
|  | CW1_GHQ128 | able to face up to problems |
|  | CW1_GHQ129 | feeling unhappy or depressed |
|  | CW1_GHQ1210 | losing confidence in self |
|  | CW1_GHQ1211 | thinking of self as a worthless person |
|  | CW1_GHQ1212 | feeling reasonably happy |
| DV_GAD | CW1_GAD2PHQ2_1 | Feeling Nervous, Anxious Or On Edge |
|  | CW1_GAD2PHQ2_2 | Not Being Able To Stop Or Control Worrying |
|  | CW1_GAD2PHQ2_3 | Little Interest Or Pleasure In Doing Things |
|  | CW1_GAD2PHQ2_4 | Feeling Down, Depressed Or Hopeless |
| DV_RPTT | CW1_RISK | willingness to take risks from 0 (Never) to 10 (Always) |
|  | CW1_PATIENT | how patient Respondent is from 0 (Never) to 10 (Always) |
|  | CW1_TRUST | how trusting of others from 0 (Not at all) to 10 (Extremely) |
|  | CW1_TRUSTPOLP | how trusting of Govmnt from 0 (Not at all) to 10 (Extremely) |

### Select rows from the Wave 1 dataset

```{r}

library(dplyr)

W01_A <- W01 %>%
  select(
    CW1_PSEX,
    CW1_COVIDSYMPT_1, CW1_COVIDSYMPT_2, CW1_COVIDSYMPT_3, CW1_COVIDSYMPT_4,
    CW1_COVIDSYMPT_5, CW1_COVIDSYMPT_6, CW1_COVIDSYMPT_7, CW1_COVIDSYMPT_8,
    CW1_COVIDSYMPT_18, CW1_COVIDSYMPT_10, CW1_COVIDSYMPT_11, CW1_COVIDSYMPT_12,
    CW1_COVIDSYMPT_16, CW1_COVIDSYMPT_13, CW1_COVIDSYMPT_14, CW1_COVIDSYMPT_17,
    CW1_COVIDSYMPT_19, CW1_COVIDSYMPT_20, CW1_COVIDSYMPT_23,
    CW1_GHQ, CW1_GHQPRECOVID,
    
    matches("^CW1_LLI_([1-9]|1[0-7])$"),
    matches("^CW1_LONELY_([1-4])$"),
    
    CW1_PHDE, CW1_PHHO, CW1_PHRF, CW1_PHEE, CW1_PHWO, CW1_PHNE,
    CW1_HARM,
    
    matches("^CW1_WEMWBS_([1-7])$"),
    matches("^CW3_MALAISE_([1-9]|1[0-9])$"),
    matches("^CW1_GHQ12([1-9]|1[0-2])$"),
    matches("^CW1_GAD2PHQ2_([1-4])$"),
    
    CW1_RISK, CW1_PATIENT, CW1_TRUST, CW1_TRUSTPOLP
  )
```

## Wave 2

**Wave 2** Variable Table

| Type       | Variable Acronym       | Explanation |
|------------|------------------------|-------------|
| Demo       | CW2_PSEX               | Sex of Respondent |
| Mod_Cov    | CW2_COVIDSYMPT_1       | Past 2 weeks COVID19 symptoms: Fever |
| Mod_Cov    | CW2_COVIDSYMPT_2       | Past 2 weeks COVID19 symptoms: Cough - dry |
| DV_WEMWBS  | CW2_WEMWBS_1           | feeling optimistic about the future |
|            | CW2_WEMWBS_2           | feeling useful |
|            | CW2_WEMWBS_3           | feeling relaxed |
|            | CW2_WEMWBS_4           | dealing with problems well |
|            | CW2_WEMWBS_5           | thinking clearly |
|            | CW2_WEMWBS_6           | feeling close to other people |
|            | CW2_WEMWBS_7           | able to make up my own mind about things |
| DV_GAD     | CW2_GAD2PHQ2_1         | Feeling Nervous, Anxious Or On Edge |
|            | CW2_GAD2PHQ2_2         | Not Being Able To Stop Or Control Worrying |
|            | CW2_GAD2PHQ2_3         | Little Interest Or Pleasure In Doing Things |
|            | CW2_GAD2PHQ2_4         | Feeling Down, Depressed Or Hopeless |
| DV_RPTT    | CW2_RISK               | willingness to take risks from 0 (Never) to 10 (Always) |
|            | CW2_PATIENT            | how patient respondent is from 0 (Never) to 10 (Always) |
|            | CW2_TRUST              | how trusting of others from 0 (Not at all) to 10 (Extremely) |
|            | CW2_TRUSTPOLP          | how trusting of Government from 0 (Not at all) to 10 (Extremely) |

```{r}

# 选择变量，自动忽略不存在的
W02_A <- W02 %>%
  select(any_of(c(
    "CW2_PSEX",
    "CW2_COVIDSYMPT_1", "CW2_COVIDSYMPT_2", "CW2_COVIDSYMPT_3", "CW2_COVIDSYMPT_4",
    "CW2_COVIDSYMPT_5", "CW2_COVIDSYMPT_6", "CW2_COVIDSYMPT_7", "CW2_COVIDSYMPT_8",
    "CW2_COVIDSYMPT_10", "CW2_COVIDSYMPT_11", "CW2_COVIDSYMPT_12", "CW2_COVIDSYMPT_13",
    "CW2_COVIDSYMPT_14", "CW2_COVIDSYMPT_16", "CW2_COVIDSYMPT_17", "CW2_COVIDSYMPT_18",
    "CW2_COVIDSYMPT_19", "CW2_COVIDSYMPT_20", "CW2_COVIDSYMPT_23",

    "CW2_PHDE", "CW2_PHHO", "CW2_PHRF", "CW2_PHEE", "CW2_PHWO", "CW2_PHNE", "CW2_HARM",

    # DV blocks
    paste0("CW2_LLI_", 1:17),
    paste0("CW2_LONELY_", 1:4),
    paste0("CW2_WEMWBS_", 1:7),
    paste0("CW2_GHQ12", 1:12),
    paste0("CW2_GAD2PHQ2_", 1:4),

    # RPTT
    "CW2_RISK", "CW2_PATIENT", "CW2_TRUST", "CW2_TRUSTPOLP"
  ))) %>%
  mutate(wave = 2)
```

```{r}
W03 <- read_sav("/Users/lisiyi/Desktop/LDAP/SiyiQiaoyu-Project/00.Data/02. covid-19_wave2_survey_cls.sav", user_na = TRUE)
```



## Wave 3

**Wave 3** Variable Table

| Type       | Variable Acronym       | Explanation |
|------------|------------------------|-------------|
| Demo       | CW3_PSEX               | Sex of Respondent |
| Mod_Cov    | CW3_COVIDSYMPT_1       | Past 2 weeks COVID19 symptoms: Fever |
| Mod_Cov    | CW3_COVIDSYMPT_2       | Past 2 weeks COVID19 symptoms: Cough - dry |
| DV_WEMWBS  | CW3_WEMWBS_1           | feeling optimistic about the future |
|            | CW3_WEMWBS_2           | feeling useful |
|            | CW3_WEMWBS_3           | feeling relaxed |
|            | CW3_WEMWBS_4           | dealing with problems well |
|            | CW3_WEMWBS_5           | thinking clearly |
|            | CW3_WEMWBS_6           | feeling close to other people |
|            | CW3_WEMWBS_7           | able to make up my own mind about things |
| DV_MALAISE | CW3_MALAISE_1          | feel tired most of the time |
|            | CW3_MALAISE_2          | feel miserable or depressed |
|            | CW3_MALAISE_3          | get worried about things |
|            | CW3_MALAISE_4          | get in a violent rage |
|            | CW3_MALAISE_5          | suddenly become scared for no good reason |
|            | CW3_MALAISE_6          | easily upset or irritated |
|            | CW3_MALAISE_7          | constantly keyed up and jittery |
|            | CW3_MALAISE_8          | every little thing gets on your nerves and wears you out |
|            | CW3_MALAISE_9          | heart often races like mad |
| DV_GAD     | CW3_GAD2PHQ2_1         | Feeling Nervous, Anxious Or On Edge |
|            | CW3_GAD2PHQ2_2         | Not Being Able To Stop Or Control Worrying |
|            | CW3_GAD2PHQ2_3         | Little Interest Or Pleasure In Doing Things |
|            | CW3_GAD2PHQ2_4         | Feeling Down, Depressed Or Hopeless |
| DV_RPTT    | CW3_RISK               | willingness to take risks from 0 (Never) to 10 (Always) |
|            | CW3_PATIENT            | how patient respondent is from 0 (Never) to 10 (Always) |
|            | CW3_TRUST              | how trusting of others from 0 (Not at all) to 10 (Extremely) |
|            | CW3_TRUSTPOLP          | how trusting of Government from 0 (Not at all) to 10 (Extremely) 


```{r}
# 选择变量，自动忽略不存在的
W03_A <- W03 %>%
  select(any_of(c(
    "CW3_PSEX",
    "CW3_COVIDSYMPT_1", "CW3_COVIDSYMPT_2", "CW3_COVIDSYMPT_3", "CW3_COVIDSYMPT_4",
    "CW3_COVIDSYMPT_5", "CW3_COVIDSYMPT_6", "CW3_COVIDSYMPT_7", "CW3_COVIDSYMPT_8",
    "CW3_COVIDSYMPT_10", "CW3_COVIDSYMPT_11", "CW3_COVIDSYMPT_12", "CW3_COVIDSYMPT_13",
    "CW3_COVIDSYMPT_14", "CW3_COVIDSYMPT_16", "CW3_COVIDSYMPT_17", "CW3_COVIDSYMPT_18",
    "CW3_COVIDSYMPT_19", "CW3_COVIDSYMPT_20", "CW3_COVIDSYMPT_23",
    
    "CW3_PHDE", "CW3_PHHO", "CW3_PHRF", "CW3_PHEE", "CW3_PHWO", "CW3_PHNE", "CW3_HARM",

    # DV blocks
    paste0("CW3_LLI_", 1:17),
    paste0("CW3_LONELY_", 1:4),
    paste0("CW3_WEMWBS_", 1:7),
    paste0("CW3_MALAISE_", 1:9),
    paste0("CW3_GAD2PHQ2_", 1:4),

    # RPTT
    "CW3_RISK", "CW3_PATIENT", "CW3_TRUST", "CW3_TRUSTPOLP"
  ))) %>%
  mutate(wave = 3)  # 标注当前数据是第3轮

```

