---
title: "Data Exploration"
format: html
editor: visual
---

## Data Exploration

Load the packages. If not installed please install first.

Overview of data

```{r}
library(foreign)
library(dplyr)


W01 <- read.spss("../00.Data/01. covid-19_wave1_survey_cls.sav", to.data.frame = TRUE)
summary(W01)
head(W01)

W02 <- read.spss("../00.Data/02. covid-19_wave2_survey_cls.sav", to.data.frame = TRUE)
summary(W02)
head(W02)

W03 <- read.spss("../00.Data/03. covid-19_wave3_survey_cls.sav", to.data.frame = TRUE)
summary(W03)
head(W03)

```

### Select Relevant Rows

The study looks into the participants with long-standing illness. Hence the following variable will be selected.

\*Demo = Demographic information Acronym

\*Mod_Cov = Moderator Covid

**Wave 1**

| Type | Variable Acronym | Explanation |
|---------|-----------------|----------------------------------------------|
| Demo | CW1_PSEX | Sex of Respondent |
| Mod_Cov | CW1_COVIDSYMPT_1 | Past 2 weeks COVID19 symptoms: Fever |
| Mod_Cov | CW1_COVIDSYMPT_2 | Past 2 weeks COVID19 symptoms: Cough - dry |
| Mod_Cov | CW1_COVIDSYMPT_3 | Past 2 weeks COVID19 symptoms: Cough - mucus or phlegm |
| Mod_Cov | CW1_COVIDSYMPT_4 | Past 2 weeks COVID19 symptoms: Sore throat |
| Mod_Cov | CW1_COVIDSYMPT_5 | Past 2 weeks COVID19 symptoms: Chest tightness |
| Mod_Cov | CW1_COVIDSYMPT_6 | Past 2 weeks COVID19 symptoms: Shortness of breath |
| Mod_Cov | CW1_COVIDSYMPT_7 | Past 2 weeks COVID19 symptoms: Runny nose |
| Mod_Cov | CW1_COVIDSYMPT_8 | Past 2 weeks COVID19 symptoms: Nasal congestion |
| Mod_Cov | CW1_COVIDSYMPT_18 | Past 2 weeks COVID19 symptoms: Sneezing |
| Mod_Cov | CW1_COVIDSYMPT_10 | Past 2 weeks COVID19 symptoms: Muscle or body aches |
| Mod_Cov | CW1_COVIDSYMPT_11 | Past 2 weeks COVID19 symptoms: Fatigue |
| Mod_Cov | CW1_COVIDSYMPT_12 | Past 2 weeks COVID19 symptoms: Unusual loose motions or diarrhoea |
| Mod_Cov | CW1_COVIDSYMPT_16 | Past 2 weeks COVID19 symptoms: Vomiting |
| Mod_Cov | CW1_COVIDSYMPT_13 | Past 2 weeks COVID19 symptoms: Loss of smell |
| Mod_Cov | CW1_COVIDSYMPT_14 | Past 2 weeks COVID19 symptoms: Loss of taste |
| Mod_Cov | CW1_COVIDSYMPT_17 | Past 2 weeks COVID19 symptoms: Skin rash |
| Mod_Cov | CW1_COVIDSYMPT_19 | Past 2 weeks COVID19 symptoms: Headaches |
| Mod_Cov | CW1_COVIDSYMPT_20 | Past 2 weeks COVID19 symptoms: Other |
| Mod_Cov | CW1_COVIDSYMPT_23 | Past 2 weeks no COVID19 symptoms experienced |
| Mod_Health | CW1_GHQ | Post-C19: Respondent self-assessment of general health |
| Mod_Health | CW1_GHQPRECOVID | Pre-C19: Respondent self-assessment of general health in 3 months preceding outbreak |
| IV | CW1_LLI_1 | long-standing illness 1: Cancer |
|  | CW1_LLI_2 | long-standing illness: Cystic fibrosis |
|  | CW1_LLI_3 | long-standing illness: Asthma |
|  | CW1_LLI_4 | long-standing illness: Chronic Obstructive Pulmonary Disease |
|  | CW1_LLI_5 | long-standing illness: Wheezy bronchitis |
|  | CW1_LLI_6 | long-standing illness: Diabetes |
|  | CW1_LLI_7 | long-standing illness: Recurrent backache/prolapsed disc/sciatica/other back probm |
|  | CW1_LLI_8 | long-standing illness: Problems with hearing |
|  | CW1_LLI_9 | long-standing illness: High blood pressure |
|  | CW1_LLI_10 | long-standing illness: Heart disease, congenital or acquired |
|  | CW1_LLI_11 | long-standing illness: Depression or other emotional/nervous/psychiatric probs |
|  | CW1_LLI_12 | long-standing illness: Obesity |
|  | CW1_LLI_13 | long-standing illness: Chronic Obstructive Airways Disease |
|  | CW1_LLI_14 | long-standing illness: Infection |
|  | CW1_LLI_15 | long-standing illness: HIV/Immunodeficiency |
|  | CW1_LLI_16 | long-standing illness: Conditions affecting the brain and nerves |
|  | CW1_LLI_17 | Respondent suffering from a long-standing illness: |
| Mod_Lonely | CW1_LONELY_1 | How often - lack companionship |
|  | CW1_LONELY_2 | How often - left out |
|  | CW1_LONELY_3 | How often - isolated from others |
|  | CW1_LONELY_4 | How often - lonely |
| DV_Kessler6 | CW1_PHDE | Felt depressed |
|  | CW1_PHHO | felt hopeless |
|  | CW1_PHRF | felt restless |
|  | CW1_PHEE | felt everything was an effort |
|  | CW1_PHWO | felt worthless |
|  | CW1_PHNE | felt nervous |
|  | CW1_HARM | has harmed self on purpose |
| DV\_ | CW1_WEMWBS_1 |  |
|  | CW1_WEMWBS_2 |  |
|  | CW1_WEMWBS_3 |  |
|  | CW1_WEMWBS_4 |  |
|  | CW1_WEMWBS_5 |  |
|  | CW1_WEMWBS_6 |  |
|  | CW1_WEMWBS_7 |  |
|  |  |  |
|  |  |  |
|  |  |  |

: select

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
